---
title: Title of your project goes here
author:
  - name: Your Name Here
    affil: 1
    orcid: '0000-0000-0000-0000'
affiliation:
  - num: 1
    address: Brown School, Washington University in St. Louis
column_numbers: 4
output: 
  posterdown::posterdown_html:
    self_contained: true
poster_height: "36in"
poster_width: "48in"
body_textsize: "36px"
sectitle_textsize: "48px"
primary_colour: "#a51417"
---

```{r setup, include=FALSE}
# Before you start make sure you have the posterdown package installed
knitr::opts_chunk$set(
  echo = FALSE,
  results = FALSE,
  error = FALSE,
  warning = FALSE,
  out.width = "90%",
  out.height = "110%",
  dpi = 400)

source(file = "project-code.R")
```

```{css, echo=FALSE}
table.Rtable1 {
    font-family: "Times New Roman", serif;
    border-collapse: collapse;
    font-size: 28pt;
    margin-left: auto;
    margin-right: auto
}
```

# Research Purpose

[[bullet point with the research question]]

# Data Description

[[bullet points describing the data]]

# Variable Measurement

Outcome:

Predictors:


# Participant Characteristics

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Table of descriptives
# add the name of the table you created in the project-code file
# in this code chunk

```

[[bullet points interpreting descriptive statistics]]

# Analysis: Outcome and Predictor 1 (replace this section title with your variables, for exampe: "Birth Control Type by Income")

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Figure for outcome + predictor 1
# add the name of the figure you created in the project-code file
# in this code chunk

```

[[bullet points to interpret the figure]]


[[bullet points to report and interpret the group statistics]]


# Analysis: Outcome and Predictor 2 (replace this section title with your variables)

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Figure for outcome + predictor 2
# add the name of the figure you created in the project-code file
# in this code chunk

```

[[bullet points to interpret the figure]]


[[bullet points to report and interpret the group statistics]]

# Analysis: Outcome and Predictor 3 (replace this section title with your variables)

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Figure for outcome + predictor 3
# add the name of the figure you created in the project-code file
# in this code chunk

```

[[bullet points to interpret the figure]]


[[bullet points to report and interpret the group statistics]]


# Summary and Recommendations

[[summarize results and add citations]]

[[write recommendations]]

# References

::: {#refs}
:::
