---
main_findings: "Increase Flu and COVID-19 Vaccination"
main_textcol: "#ffffff"
column_numbers: 2
output: 
  posterdown::posterdown_betterport:
    self_contained: true
poster_height: "11in"
poster_width: "8.5in"
body_textsize: "14px"
main_textsize: "54px"
sectitle_textsize: "16px"
primary_colour: "#3182bd"
secondary_colour:	"#3182bd"
bibliography: references.bib
knit: pagedown::chrome_print
---

```{r setup, include=FALSE}
#| echo: false
#| results: false
#| warning: false
#| error: false
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  out.width = "90%",
  out.height = "110%",
  dpi = 400)

source(file = "richards-course-project-code.R")
```

## Flu & COVID-19 Vaccination Intention

* Flu and COVID-19 vaccines are effective against serious disease and reduce transmission. 

* Fewer people intend to get vaccinated against COVID-19 compared to flu.

* Older people are less likely to intend to get a COVID-19 vaccine.

* More conservative people are less likely to intend to get a COVID-19 vaccine.

```{r echo=FALSE, message=FALSE, warning=FALSE}
vaxMapFlu
intentionMapCovid
```

## More Flu Vaccination -> More Covid Vaccination

* **89%** of people who had the flu shot intended to get the COVID-19 vaccine.

```{r}
waffleplot
```

## What Can We Do? 

* Increasing the uptake of flu vaccination and developing health communication campaigns that reach older and more conservative audiences could increase the uptake of COVID-19 vaccination.

## Contact Info & Sources

* Contact Me: 
  + Megan Richard at megan.r\@wustl.edu
  
* Sources:
  + CDC COVID Data Tracker and FluVaxView
  + Research Project: *Motivations and Barriers to Vaccination during Covid-19.*
  + Albrecht, 2022. *Vaccination, Politics and COVID-19 Impacts.*
  + Rasul and Saifuddin, 2023. *Not All Conservatives Are Vaccine Hesitant.*
  + Ruiz and Bell, 2021. *Predictors of Intention to Vaccinate Against COVID-19.*
