---
title: Motivations and Barriers to Vaccination during Covid-19
author:
  - name: Megan Richard
    affil: 1
    orcid: '0000-0000-0000-0000'
affiliation:
  - num: 1
    address: Brown School, Washington University in St. Louis
column_numbers: 4
output: 
  posterdown::posterdown_html:
    self_contained: true
poster_height: "36in"
poster_width: "48in"
body_textsize: "36px"
sectitle_textsize: "48px"
primary_colour: "#a51417"
bibliography: references.bib
knit: pagedown::chrome_print
---

```{r setup, include=FALSE}
#| echo: false
#| results: false
#| warning: false
#| error: false
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  out.width = "90%",
  out.height = "110%",
  dpi = 400)

source(file = "richards-course-project-code.R")
```

```{css, echo=FALSE}
table.Rtable1 {
    font-family: "Times New Roman", serif;
    border-collapse: collapse;
    font-size: 28pt;
    margin-left: auto;
    margin-right: auto
}
```

# Research Purpose

-   This research seeks to answer the following question: Are age, political beliefs, and prior flu shot associated with the intention to get the COVID-19 vaccine for adults in the United States?

# Data Description

-   Data were collected in October 2020 through a Google Survey disseminated on social media

-   Survey questions were chosen by evaluating existing literature [@rasul2023; @ruiz2021; @albrecht2022]

-   After removing observations missing data on the 4 relevant variables, 4 variables and 511 observations will be examined for this project

-   The codebook can be downloaded from [this link](https://github.com/OxCGRT/covid-policy-tracker/blob/master/documentation/codebook.md)

# Variable Measurement

Outcome:

-   Intention to get the COVID-19 vaccine (categorical): Yes, No, Don't Know/Not Sure

Predictors:

-   Political beliefs (categorical): Very Liberal, Somewhat Liberal, Mixed, Somewhat Conservative, Very Conservative

-   Flu shot in the past 12 months (categorical); Yes, No

-   Age in years (continuous)

# Participant Characteristics

```{r echo=FALSE, warning=FALSE, message=FALSE}
# print table
table1


```

-   Nearly half of participants intended to get the covid vaccine (47.5%; n = 242)

-   Most survey participants had received a flu shot in the past year (64.6%; n = 330)

-   Most participants were somewhat or very liberal (55.3%; n = 282)

-   The median age of participants was 26 years old (IQR = 25)

# Vaccine Intention and Flu Shot History

-   Figure 1 shows a higher percentage of people who intend to get the COVID-19 vaccine have had the flu shot and the opposite is true for those not intending to get the COVID-19 vaccine

    -   Of those who intend to get the COVID-19 vaccine, 79.3% of them had a flu shot in the last year and 20.7% did not have a flu shot

    -   Of those who do not intend to get the COVID-19 vaccine, only 37.5% of them had a recent flu shot and 62.5% did not have a flu shot

```{r echo=FALSE, message=FALSE, warning=FALSE}
figure1
```



-   Flu shot history and vaccine intention are both categorical so I used chi-squared to determine whether they were statistically significantly associated

    -   All chi-squared assumptions were met including independent observations, both variables were nominal or ordinal, and more than 80% of expected values were greater than 5

-   There is a statistically significant association between vaccine intention and flu shot history ($\chi^2$ = 56.35; p \< .05)

-   Standardized residuals showed which groups were driving the statistically significant result

    -   Significantly **fewer than expected** people who had a flu shot did not intend to take the COVID-19 vaccine (std res = -3.30) and **more than expected** who had a flu shot intended to take the vaccine (std res = 2.87)

    -   Significantly **more than expected** people with no flu shot did not intend to take the COVID-19 vaccine (std res = 4.44) and **fewer than expected** with no flu shot intended to take the vaccine (std res = -3.87)

# Vaccine Intention and Political Beliefs

-   Figure 3 suggests that political beliefs may be associated with intention to get the COVID-19 vaccine

    -   Of those who intend to get the COVID-19 vaccine, 50% were very liberal and 24.8% were somewhat liberal

    -   Of those who do not intend to get the COVID-19 vaccine, 32.3% were politically mixed and 21.9% were somewhat conservative

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=7}
figure3
```

-   Political beliefs and vaccine intention are both categorical so I used chi-squared to determine whether they were statistically significantly associated

    -   All chi-squared assumptions were met including independent observations, both variables were nominal or ordinal, and more than 80% of expected values are greater than 5

-   I found a statistically significant association between vaccine intention and political beliefs ($\chi^2$ = 142.55; p \< .05).

-   Standardized residuals showed which groups were driving the statistically significant result

    -   Significantly **more very liberal people than expected were intending to take the COVID-19 vaccine** (std res = 4.83) while **significantly fewer than expected** who preferred not to answer, were politically mixed, or were somewhat conservative were intending to take the vaccine (std res \< -2)

    -   Significantly **more than expected** people who preferred not to answer, were politically mixed, or were very conservative were not intending to take the COVID-19 vaccine (std res \> 2) while **significantly fewer than expected** who were very liberal or somewhat liberal were intending to not take the vaccine (std res \< -2)

# Vaccine Intention and Age

-   The median age is highest in the "No" vaccine intention group (median = 39.5, IQR = 23.8--50.0) and lowest in the "Yes" vaccine intention group (median = 24.0, IQR = 22.0--41.8)

-   The large difference suggests that age is likely associated with vaccine intention and that there are more older participants who do not intend to get the COVID-19 vaccine

```{r echo=FALSE, message=FALSE, warning=FALSE}
figure4
```

-   The ANOVA assumption of normal distribution within groups was not met with the distribution of age being very right skewed in the group who intended to get the COVID-19 vaccine

-   Age is continuous and vaccine intention is categorical with 3 categories so I used the Kruskal-Wallis test to test whether the distribution of ranked age was the same across vaccine intention groups

-   There was a statistically significant difference in age distribution across the vaccine intention groups (KW $\chi^2$ = 11.63; p = .003)

-   A post-hoc Dunn's test can determine which groups are statistically significantly different from each other

    -   The mean age of the group with no intention to vaccinate for COVID-19 was statistically significantly higher (diff = 3.35 years) compared to the group that intended to vaccinate for COVID-19 (p = .001)

# Summary and Recommendations

Age, having a recent flu vaccine, and political beliefs were all statistically significantly associated with intent to get a COVID-19 vaccination. Consistent with prior studies, I found that people who had a flu vaccination in the last year, had more liberal political beliefs, and were younger were more likely to intend to get the COVID-19 vaccine [@rasul2023; @ruiz2021; @albrecht2022].

Increasing the uptake of flu vaccination and developing health communication campaigns that reach older and more conservative audiences could both increase the uptake of COVID-19 vaccination.

# References

::: {#refs}
:::
